---
title: "Class 6 - Tables and Plots"
author: "Or Duek"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: ioslides_presentation
---

## Today we'll build tables and plots

- R has some functions for representing tables and description
- Most common are *table* (R base), *table1* and *gtsummary*

## Table

* Let's grab a dataframe and show something
```{r message=FALSE}
# we first load packages
library(readxl)
library(tidyverse)
library(janitor)
# we then read the data
df <- read.csv('Weight and Height.csv')
df <- df %>% clean_names()
```

## Show the data in table
```{r}
table(df$what_is_you_gender, df$what_is_your_weight_in_kg)
```

## Not the best
Let use something better
```{r message=FALSE}
library(table1)
table1(~ what_is_your_height_in_cm | what_is_you_gender, data = df)

```

## Lets add height and weight
```{r}
table1( ~ what_is_your_height_in_cm  + what_is_your_weight_in_kg | what_is_you_gender, data=df)
```

# COOL!!

## Lets use gtsummary now

```{r message=FALSE}
library(gtsummary)
gtsummary::theme_gtsummary_journal("nejm")

df |> select(what_is_your_height_in_cm, what_is_your_weight_in_kg, what_is_you_gender) |>   tbl_summary(by = what_is_you_gender,
  label = list(what_is_your_height_in_cm = "Height (cm)", what_is_your_weight_in_kg = "Weight (Kg)")) |> add_p() |> add_overall()
```

# Lets do some Graphs!!

## Graphs
* Basic include
  - plot(x,y)
  - hist(x)
  - pie(x)
  - barplot(y, names.arg = x)

## Use basic
```{r}
plot(df$what_is_your_height_in_cm, df$what_is_your_weight_in_kg)
```

## Bar

```{r}
barplot(df$what_is_your_height_in_cm)
```


# Introducing *ggplot2*


## Used a lot
* *ggplot2* is used many many times
* you can check the link here [https://exts.ggplot2.tidyverse.org/gallery/]


## Using it
```{r message=FALSE}
library(ggplot2)

ggplot(df, aes(x=what_is_you_gender, y=what_is_your_height_in_cm, 
               fill=what_is_you_gender)) + 
  geom_boxplot() + theme_minimal()
               
```


## Using it
```{r}
ggplot(df, aes(x=what_is_your_weight_in_kg, y=what_is_your_height_in_cm
               , color = what_is_you_gender)) + 
  geom_point() + theme_minimal()
```

## Adding line?
```{r}
ggplot(df, aes(x=what_is_your_weight_in_kg, y=what_is_your_height_in_cm,
               color=what_is_you_gender)) + 
  geom_point() + geom_smooth(method= 'lm', alpha = .1) + theme_minimal() +
  labs(x="Kg")
```




## Reshaping Data Frames

```{r}
#creating a wide dataframe
df_wide <- data.frame(
  Name = c("Alice", "Bob"),
  Score_2021 = c(88, 92),
  Score_2022 = c(91, 95)
)

print(df_wide)
```

## Wide to Long Transformation
Using `pivot_longer()` to transform data from wide format to long format.

```{r}
df_long<- df_wide %>% pivot_longer(cols=c("Score_2021","Score_2022"))
df_long
```

## Setting new variable names
```{r}
df_long <- df_wide %>% pivot_longer(cols=c("Score_2021","Score_2022"),
                                  names_to = 'year', values_to = 'score')
df_long
```

## One last tyding
lets remove this Score_ from the data
```{r}
df_long <- df_wide %>% pivot_longer(cols=c("Score_2021","Score_2022"),
                                  names_to = 'year', values_to = 'score',
                                  names_prefix = 'Score_')
df_long
```

### Long to Wide Transformation
Using `pivot_wider()` to transform data from long format to wide format.
```{r}
dfw <- df_long %>% pivot_wider(names_from = year, 
                               values_from = score)

dfw
```
