---
title: "Functions and Loops"
author: "Or Duek"
date: "`r Sys.Date()`"
output: html_document
---

# In this class we will learn how to build and use R function and how to loop 
# through


# Building you own function
- Lets say we want to calculate BMI, or we want to calculate the next number that follows (so if I get the number 1, i'll return the number 2 etc.)
- We can build our own function to do so
- How?

let's see

```{r}
nextNum <- function(x) {
  y <-  x + 1
  return (y)
}
```

## The function is built from its name (nextNum), then a call for a function and then the input it receives (x in this case)
We then do a number of actions on x and return whatever we want (in this case, y)

- Lets use it
```{r}
nextNum(5)
```

Now, go and build your own function. 
- Make a function that receives weight and height and returns BMI
- Remember, $BMI = \frac{height(m)}{weight^2}$

```{r}
# make your own function
```


# Loops

I want to show you how loops work
Then we will present an easier way to do so

- For loops
  - for loops basically runs i number of times, each time they do something
  
```{r}
for(i in 1:10) {
  print(i)
}
```
# across() function in dplyr
- *across()* allows us to run a function on multiple columns

- let's try it
```{r}
library(dplyr)
library(janitor)
library(ggplot2)
df <- read.csv('students_depression_data.csv')
df <- clean_names(df)
names(df)
```
# let's do average for scaling (z score) for a few numeric variables here

- we srart simply by scaling the following

```{r}
df %>% 
  mutate(across(c(age, academic_pressure, cgpa), scale))

```

## You can see that the original variables are now scaled
- But, if we want to keep the original variabels, we can use the *.names* argument
- Within the *.names* argument we can use {.col} and {.fn} to describe the column original name and the function used. 

Let's see how it works:

```{r}
df %>% 
  mutate(across(c(age, academic_pressure, cgpa), scale, .names="{.col}_scaled"))
```

# COOL!!

Now, let plot these two:

```{r}
df1 <- df %>% 
  mutate(across(c(age, academic_pressure, cgpa), scale, .names="{.col}_scaled"))

ggplot(df1, aes(x=cgpa)) + geom_histogram() + theme_minimal() + labs(title = "Unscaled CGPA")

ggplot(df1, aes(x=cgpa_scaled)) + geom_histogram() + theme_minimal() + labs(title = "Scaled CGPA")
```


