---
title: "Students Depression"
author: "Or Duek"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(janitor)
library(gtsummary)
```

## Student's Depression Dataset

We now load and start to play with Student's depression dataset

```{r read_data}
df <- read.csv('students_depression_data.csv')
```

## Including Plots

You can also embed plots, for example:

```{r}
summary(df)
```
We can see that gender is a character - we want it to be a factor (also profession, degree, dietary habits, suicidal thoughts)
We can also see names of variables are not cool


```{r}
df <- janitor::clean_names(df)
summary(df)
```

```{r }
df <- df %>% mutate(
  profession = as.factor(profession),
  degree  = as.factor(degree),
  have_you_ever_had_suicidal_thoughts = as.factor(have_you_ever_had_suicidal_thoughts),
  gender = as.factor(gender), 
  dietary_habits = as.factor(dietary_habits),
  family_history_of_mental_illness = as.factor(family_history_of_mental_illness)
  
)
```


Ok, lets see now

```{r}
summary(df)
```

## Lets do table of gender and depression

```{r echo=FALSE}
df %>% select(gender, depression) %>%
  tbl_summary(
    by = gender
  ) %>% add_p()
```

Ohh - depression is 0 or 1 but R read it as a numeric variable. We need to change it

```{r}
df <- df %>% mutate(
  depression = as.factor(depression)
)
```

# Now lets plot it?

```{r}
ggplot(df, aes(x=gender, y=depression, color = gender)) + geom_count() + 
     theme_minimal()
```

## Ok(ish), lets try something else
```{r}
df %>%
  dplyr::count(gender, depression) %>% ggplot(aes(x = gender, y = n, fill = depression)) +
  geom_bar(stat="identity", position = "dodge") +
  labs(title = "Depression Status by Gender",
       x = "Gender",
       y = "Count",
       fill = "Depression") +
  theme_minimal()
```
### Lets make it stacked
```{r}
df %>%
  dplyr::count(gender, depression) %>% ggplot(aes(x = depression, y = n, fill = gender)) +
  geom_bar(stat = "identity", position = "stack") + scale_fill_manual(values = c("Male" = "#1f77b4", "Female" = "#ff7f0e")) +
  labs(title = "Depression Status by Gender",
       x = "Depression",
       y = "Count",
       fill = "Gender") + 
  theme_minimal()
```

## Let's test if there is a difference between Gender in financial stress


```{r}
ggplot(df, aes(x=gender, y = financial_stress, fill= gender)) + geom_boxplot() +
    theme_minimal()
```

Now we run a t-test
```{r}
t.test(financial_stress ~ gender, data=df)

```
## Let's test GPA and sleep
```{r}
ggplot(df, aes(x=sleep_duration, y=cgpa, fill=sleep_duration, color = sleep_duration)) + geom_jitter(alpha=.01) + geom_boxplot() +
  theme_minimal()
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
