---
title: "Class 7 - Wide and Long (and some more plotting)"
author: "Or Duek"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: ioslides_presentation
---

# Today we'll pivot long and wide

## First load libraries

```{r message=FALSE}
# we first load packages
library(readxl)
library(tidyverse)
library(janitor)
```


## Reshaping Data Frames

* We generate a fake data frame

```{r}
#creating a wide dataframe
df_wide <- data.frame(
  Name = c("Alice", "Bob"),
  Sex = c("Female","Male"),
  Score_2021 = c(88, 92),
  Score_2022 = c(91, 95)
)

print(df_wide)
```

## Wide to Long Transformation
Using `pivot_longer()` to transform data from wide format to long format.

```{r}
df_long<- df_wide %>% pivot_longer(cols=c("Score_2021","Score_2022"))
df_long
```

## Setting new variable names
```{r}
df_long <- df_wide %>% pivot_longer(cols=c("Score_2021","Score_2022"),
                                  names_to = 'year', values_to = 'score')
df_long
```

## One last tyding
lets remove this Score_ from the data
```{r}
df_long <- df_wide %>% pivot_longer(cols=c("Score_2021","Score_2022"),
                                  names_to = 'year', values_to = 'score',
                                  names_prefix = 'Score_')
df_long
```


## Long to Wide Transformation

Using `pivot_wider()` to transform data from long format to wide format.

```{r}
dfw <- df_long %>% pivot_wider(names_from = year, 
                               values_from = score)

dfw
```

```{r}
df_age <- read_csv('age.csv')

```

