---
title: "Lecture 3 - Variables in R"
author: "Or Duek"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  ioslides_presentation: default
  beamer_presentation:
    theme: Madrid
    colortheme: dolphin
---

# Introduction to Variables in R

## What are variables?
Variables in R are objects (typically vectors) that store data. They can have different types depending on the nature of the data they store.


# Basic Variables in R

## Discrete Variable

A discrete variable is a type of numerical variable that has a countable number of values between any two values.

## Logical Variable

Logical variables are binary, indicating `TRUE` or `FALSE` states.

## Factor (Ordered Variable)

Factors are the data objects in R which are used to categorize the data and store it as levels. They can store both strings and integers.


# Useful Functions for Variables

## Conditional Assignment

```{r}
# Example of conditional assignment using ifelse
x <- c(1, 2, 3, 4, 5)
y <- ifelse(x > 3, TRUE, FALSE)
```

## Subsetting with Conditions

```{r}
x <- c(NA,2,3,NA,5)
x[is.na(x)] <- 0 # replace NA with 0
```

## Variables in Data Frames
Accessing and manipulating data frame variables

```{r}
df <- data.frame(y = 1:5, z = 6:10)
df$x <- df$y + df$z  # Create a new variable 'x' as the sum of 'y' and 'z'
df$x[is.na(df$x)] <- 0  # Change NA values in 'x' to 0
df$g <- ifelse(df$x >11, "high", "low")
```

## Common Errors
```{r}
z <- c(NA,2,3,NA,5)
mean(z)#, na.rm = TRUE)

```
## Calculating BMI
```{r}
df_hw <- read.csv('Weight and Height.csv')
df_hw$BMI <- df_hw$What.is.your.weight..in.Kg.. / ((df_hw$What.is.your.height..in.CM../100)^2)
```

## Piepelines

We can do something with multiple functions or we can use pipes (|> or %>%)

```{r}
v <- c(2,3,5,6)
mean(v)

v |> mean()
```
## Tydiverse
```{r}
#v %>% mean()
# error lets load
library(tidyverse)
v %>% mean()
```

