
---
title: "Lecture 2 - R Objects"
author: "Or Duek, Ph.D"
format: html
output:
  beamer_presentation: default
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to R Objects

R operates by using objects that store data and results of computations. Users create these objects to contain information such as data, results of mathematical operations, and statistical tests.

## Vector Creation

### Creating Vectors

You can create different types of vectors in R using the `c()` and `vector()` functions. Here are examples for each type:

- Numeric vectors:
```{r}
x <- 3
numeric_vector <- c(1, 2, 3, 4)
  ```

- Character vectors:
  ```{r}
  character_vector <- c("one", "two", "three")
  ```

- Logical vectors:
  ```{r}
  logical_vector <- c(TRUE, FALSE, TRUE)
  ```

### Accessing Vector Elements

You can access specific elements of a vector using square brackets. For example, to access the third element of the `numeric_vector`:
  ```{r}
  third_element <- numeric_vector[3]
  ```

## Matrix Creation

### Creating Matrices

You can create matrices using the `matrix()` function for two-dimensional matrices or `array()` for multi-dimensional arrays.

- Example of a numeric matrix:
  ```{r}
  matrix_example <- matrix(1:9, nrow = 3, ncol = 3)
  ```

### Accessing Matrix Elements

To access specific elements, use the row and column indices:
  ```{r}
  element <- matrix_example[3, 2]  # Third row, second column
  ```

## Data Frames

### Creating and Manipulating Data Frames

Data frames are essential in R for storing data tables. Here’s how to create a data frame:
  ```{r}
  df <- data.frame(
    Column1 = c(1, 2, 3),
    Column2 = c("A", "B", "C")
  )
  ```

### Importing Data

- Reading a CSV file:
  ```{r}
  data <- 
  read.csv("~/Documents/Untitled form.csv")
  ```

## External Data File Handling

### Reading and Writing Data

- To read an Excel file:
  ```{r}
  library(readxl)
 # excel_data <- read_excel("path/to/your/file.xlsx")
  ```

- To write data to a CSV file:
  ```{r}
  #write.csv(data, "path/to/your/output.csv")
  ```

## R Workspace Management

### Managing the Workspace

To delete an object:
  ```{r}
  rm(list = c("data", "df"))
  ```

### Saving Work Sessions

To save all workspace objects:
  ```{r}
  save.image(file = "my_workspace.RData")
  ```

To save a selected list of objects:
  ```{r}
  save(list = c("data", "df"), file = "selected_workspace.RData")
  ```
